---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# psidread

<!-- badges: start -->

<!-- badges: end -->

The goal of psidread is to provide a user-friendly approach to streamline the management, creation, and formatting of panel data from the Panel Study of Income Dynamics (PSID).

## Installation

You can install the development version of psidread from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Qcrates/psidread")
```

## Usage

### Overview

The [Panel Study of Income Dynamics](https://psidonline.isr.umich.edu/) (PSID) is the longest running longitudinal household survey in the world, which provides invaluable data covering numerous topics including marriage, income, wealth, health and etc. However, the process of converting raw PSID data files into datasets ready for analysis is quite complex and challenging, especially for new users.

This package is developed with the purpose of addressing these challenges within only R environment without additional assistance from other statistical programming softwares. By bridging these gaps, the package aims to make PSID datasets more usable and manageable for researchers and analysts.

### Introduction

What `psidread` package is created to help:

-   Create a table of data structure across multiple waves using the text that can be copied and pasted from the website

-   Unzip and convert the zipped files without additional help of other software

-   Read and merge the data files from multiple waves

-   Rename and reshape the dataset to fit the need for advanced analysis

### Example Workflow

```{r, eval = FALSE}
psid_varlist = c(" hh_age || [13]ER53017 [17]ER66017", " p_age || [13]ER34204")
str_df <- psid_str(varlist = psid_varlist, type = "separated")
input_directory <- "/Users/RachelChiu/Documents/ReProj/PSID/psidread/inst/extdata"
psid_df <- psid_read(indir = input_directory, str_df = str_df,idvars = c("ER30000"),type = "package",filename = NA)
str(psid_df)
```

Please refer to the vignettes for the detailed instructions on how to build your own dataset from PSID using this package: [psidread user manual](vignettes/my-vignette.html) (GitHub does not show the html file directly. Download and open the file in the web browser.)
